---
title: "Seabirds Analysis"
output: html_notebook
---

# Seabirds analysis

```{r}
library(tidyverse)
library(here)

seabirds_clean <- read_csv(here("clean_data/seabirds_clean.csv"))
```
## General Observations
```{r}
seabirds_clean
```

```{r}
#hide chunk
seabirds_clean %>% 
  count(species_scientific_name)
```
 - make list
There are 150 different common names.
There are 125 different scientific names, these encompass 47 families.
There are 142 different abreviations.



## Which bird had the most individual sightings?
```{r}
seabirds_clean %>% 
  group_by(species_scientific_name) %>%
  summarise(total_sightings = n()) %>% 
  arrange(desc(total_sightings))

seabirds_clean %>% 
  group_by(family) %>%
  summarise(total_sightings = n()) %>% 
  arrange(desc(total_sightings))

seabirds_clean %>% 
  select(species_common_name, species_scientific_name, species_abbreviation) %>% 
  filter(species_scientific_name == "Diomedea antipodensis") %>% 
  distinct()

```
The most common species sighted was a Wandering albatross sensu lato (Diomedea antipodensis, DIOANTEXU) at 11,293 sightings. 

## Which bird had the highest total count?
```{r}
seabirds_clean %>%
  group_by(species_scientific_name) %>% 
  summarise(total_count = sum(count, na.rm = TRUE)) %>% 
  arrange(desc(total_count))

seabirds_clean %>% 
  select(species_common_name, species_scientific_name, species_abbreviation) %>% 
  filter(species_scientific_name == "Puffinus tenuirostris") %>% 
  distinct()
```
The short-tailed shearwater (or sooty shearwater,	Puffinus tenuirostris, PUFTEN/PUFTENGRI) has the highest total count across all sightings at 996,082.

## Which bird had the highest total count above a latitude of -30?
```{r}
seabirds_clean %>%
  filter(lat > -30) %>% 
  group_by(species_scientific_name) %>% 
  summarise(total_count = sum(count, na.rm = TRUE)) %>% 
  arrange(desc(total_count))

seabirds_clean %>% 
  select(species_common_name, species_scientific_name, species_abbreviation) %>% 
  filter(species_scientific_name == "Puffinus pacificus") %>% 
  distinct()
```
The Wedge-tailed shearwater (Puffinus pacificus, PUFPAC)......GO BACK AND PRINT RESULT NICELY

## How many different types of birds were only ever seen in groups of 1?
```{r}
seabirds_clean %>% 
  group_by(species_scientific_name) %>% 
  filter(mean(count, na.rm = TRUE) == 1) %>% #this assumes that there are no species that happen to have an average of 1 despite some variation within count
  summarise()
```
There were 21 different species which were only ever seen on their own.

## How many penguins were seen?
```{r}
total_num_of_penguins <- seabirds_clean %>% 
  filter(str_detect(species_common_name, "penguin")) %>% 
  #group_by(species_common_name) %>% 
  summarise(total = n()) %>% 
  pull()

penguin_species <- seabirds_clean %>% 
  filter(str_detect(species_common_name, "penguin")) %>% 
  group_by(species_common_name) %>% 
  summarise(total = n()) %>% 
  summarise(n()) %>% 
  pull()

paste("There were", total_num_of_penguins,
      "penguins seen in total across", penguin_species, "different species.")
```



