---
title: "Seabirds Analysis"
output: html_notebook
---

# Seabirds analysis

```{r, message=FALSE}
library(tidyverse)
library(here)

seabirds_clean <- read_csv(here("clean_data/seabirds_clean.csv"))
```
## General Observations
```{r}
seabirds_clean
```

```{r, echo = FALSE}
seabirds_clean %>% 
  count(species_scientific_name)

seabirds_clean %>% 
  count(species_common_name)

seabirds_clean %>% 
  count(species_abbreviation)

seabirds_clean %>% 
  count(family)
```
* There are 47,918 rows of data and 12 columns
* The data is all in the expected format
* There are 163 different scientific names, these encompass 48 families.
* There are 152 different abbreviations for species.
* There are 159 different common names.

***

## Which bird had the most individual sightings?
```{r}
seabirds_clean %>% 
  group_by(species_scientific_name) %>%
  summarise(total_sightings = n()) %>% 
  slice_max(total_sightings, n = 1)

seabirds_clean %>% 
  select(species_common_name, species_scientific_name, species_abbreviation) %>% 
  filter(species_scientific_name == "Diomedea antipodensis") %>% 
  distinct()

```
**MOST COMMON BIRD SIGHTINGS**
Common Name                    | Species Name          | Species Abbreviation | Number of sightings
------------------------------ |---------------------- |--------------------- |-------------------- 
Wandering albatross sensu lato | Diomedea antipodensis | DIOANTEXU            | 11,293


## Which bird had the highest total count?
```{r}
seabirds_clean %>%
  group_by(species_scientific_name) %>% 
  summarise(total_count = sum(count, na.rm = TRUE)) %>% 
  arrange(desc(total_count))

seabirds_clean %>% 
  select(species_common_name, species_scientific_name, species_abbreviation) %>% 
  filter(species_scientific_name == "Puffinus tenuirostris") %>% 
  distinct()
```
The short-tailed shearwater (or sooty shearwater,	Puffinus tenuirostris, PUFTEN/PUFTENGRI) has the highest total count across all sightings at 996,082.

## Which bird had the highest total count above a latitude of -30?
```{r}
seabirds_clean %>%
  filter(lat > -30) %>% 
  group_by(species_scientific_name) %>% 
  summarise(total_count = sum(count, na.rm = TRUE)) %>% 
  arrange(desc(total_count))

seabirds_clean %>% 
  select(species_common_name, species_scientific_name, species_abbreviation) %>% 
  filter(species_scientific_name == "Puffinus pacificus") %>% 
  distinct()
```
The Wedge-tailed shearwater (Puffinus pacificus, PUFPAC) had the highest total count from sightings above a latitude of -30.

## How many different types of birds were only ever seen in groups of 1?
```{r}
seabirds_clean %>% 
  group_by(species_scientific_name) %>% 
  filter(mean(count, na.rm = TRUE) == 1) %>% #this assumes that there are no species that happen to have an average of 1 despite some variation within count
  summarise()
```
There were 21 different species which were only ever seen on their own.

## How many penguins were seen?
```{r}
total_num_of_penguins <- seabirds_clean %>% 
  filter(str_detect(species_common_name, "penguin")) %>% 
  #group_by(species_common_name) %>% 
  summarise(total = sum(count, na.rm = TRUE)) %>% 
  pull()

penguin_species <- seabirds_clean %>% 
  filter(str_detect(species_common_name, "penguin")) %>% 
  group_by(species_common_name) %>% 
  summarise(total = n()) %>% 
  summarise(n()) %>% 
  pull()

paste("There were", total_num_of_penguins,
      "penguins seen in total across", penguin_species, "different species.")
```

```{r}
seabirds_clean %>% 
  filter(!str_detect(species_common_name, "unidentified|\\/")) %>% 
  group_by(species_common_name) %>% 
  summarise(total = n()) %>% 
  arrange(desc(total)) %>% 
  head(10) %>% 
  ggplot() +
  geom_col(aes(reorder(species_common_name,total), total)) +
  coord_flip() +
  labs(
    x = "Species",
    y = "Total Number of Sightings Recorded",
    title = "Most Common Species Sighted"
  ) +
  scale_y_continuous(labels = scales::comma)
```



