---
title: "Dog Data Analysis"
author: "Emma Menzies"
output: html_notebook
---
# Set up
```{r}
library(tidyverse)
library(here)

dog_data <- read_csv(here("clean_data/dog_data_clean"))
```

```{r}
dog_data
```
## Survey Responses Received by size and gender of dog
```{r}
dog_data %>% 
  filter(!is.na(dog_size) & !is.na(dog_gender)) %>% 
  mutate(dog_size = factor(dog_size, levels = c("XS", "S", "M", "L", "XL"))) %>% 
  group_by(dog_size, dog_gender) %>% 
  summarise(total_responses = n()) %>% 
  ggplot(aes(dog_size, total_responses, fill = dog_gender)) +
  geom_col() +
  theme_classic() +
  #scale_fill_brewer(palette = "Pastel1") +
  labs(
    title = "Number of Responses Received by Size of Dog",
    x = "Size of Dog",
    y = "Number of Survey Responses",
    fill = "Gender"
  )

ggsave("responses_recieved.png")
```

# 1. Valid Email Addresses
The client only counts a valid email address as one ending in ‘.com’. How many survey results have a valid email address.
Since some rows have been split into multiple rows with the same individuals data count only distinct ids.
```{r}
dog_data %>% 
  filter(str_detect(email, ".com$")) %>% 
  summarise(total = n_distinct(id))
```

# 2. Average Dog Food Spend
What’s the average amount spent on dog food for each dog size.
```{r}
dog_data %>% 
  group_by(dog_size) %>% 
  summarise(average_spend =
              round(mean(amount_spent_on_dog_food, na.rm = TRUE), 2)) %>% 
  arrange(average_spend)
```
## Visualisation
```{r}
dog_data %>% 
  filter(!is.na(dog_size)) %>% 
  mutate(dog_size = factor(dog_size, levels = c("XS", "S", "M", "L", "XL"))) %>% 
  group_by(dog_size) %>% 
  summarise(average_spend =
              round(mean(amount_spent_on_dog_food, na.rm = TRUE), 2)) %>% 
  arrange(average_spend) %>% 
  ggplot(aes(dog_size, average_spend, fill = dog_size)) +
  geom_col(show.legend = FALSE) +
  theme_classic() +
  scale_fill_brewer(palette = "Pastel1") +
  labs(
    title = "Average Spend on Dog Food by Size",
    x = "Size of Dog",
    y = "Average Spend"
  )

ggsave("spend_vs_size.png")
```

# 3. Average age
For owners whose surname starts with a letter in the second half of the alphabet (N onwards) what is the average age of their dog?
```{r}
dog_data %>% 
  filter(str_detect(last_name, "^[N-Z]+")) %>% 
  summarise(average_age = mean(dog_age, na.rm = TRUE))
```

# 4. Dog years to human years
The dog_age column is the age in dog years. If the conversion is 1 human year = 6 dog years, then what is the average human age for dogs of each gender?
```{r}
dog_human_age <- dog_data %>% 
  mutate(dog_gender = case_match(
    dog_gender,
    "F" ~ "Female",
    "M" ~ "Male",
    NA ~ "Unknown"
  )) %>%  
  group_by(dog_gender) %>%
  summarise(average_age_in_human_years = mean(dog_age, na.rm = TRUE)/6) 

dog_human_age
```

## Visualisation
```{r}
ggplot(dog_human_age) +
  aes(dog_gender, average_age_in_human_years, fill = dog_gender) + 
  geom_col() +
  labs(
    x = "Gender",
    y = "Average age in human years",
    title = "Average dog ages converted to human years by gender"
  ) +
  scale_fill_manual(
    values = c(
      "Female" = "cadetblue",
      "Male" = "chartreuse3",
      "Unknown" = "bisque4"
    ))

ggsave("dog_ages.png")
```



