---
title: "Halloween Candy Analysis ðŸ¬"
author: "Emma Menzies"
output:
  html_document:
    df_print: paged
---

# Contents  

1. [Introduction and set up]  
2. [Question 1](#Question_1)
3. [Question 2](#Question_2)
4. [Question 3](#Question_3)
5. [Question 4](#Question_4)
6. [Question 5](#Question_5)
7. [Function](#Function)
8. [Question 6](#Question_6)
9. [Question 7](#Question_7)
10. [Question 8](#Question_8)
11. [Extra](#Extra)

# Introduction and set up

**Data**  

This analysis looks at halloween survey results from several countries over the years 2015-2017.  
Data on individual's age, gender, and country were collected along with whether they are going trick-or-treating that year.  
Individuals were also asked to rate many different candy types on a scale of joy > meh > despair.  
A range of other questions were asked which have been removed from the dataset.  

**Assumptions**  

* Where country was "Trumpistan" the country is US    
* Where country was "Cascadia" this is in "Other" as it is a real area while not being a country     
* For gender "I'd rather not say", "Other", and "Not gathered" all have different implications and have been treated separately.   
* The first numbers in the age column is the person's age   

**Cleaning** 

* A unique identifier was added for each survey response within each year: 
  + eg. ...`mutate("id" = paste0("2015_", row_number()))`... -> 2015_1, 2015_2, etc.   

* For each year of data the relevant columns were then extracted, renamed, and pivoted into the format:  
|id | year | gender | country | age | trick_or_treating | candy_type | rating | rating_score  
*Where rating_score is 1 = joy, 0 = meh, -1 = despair.*  

* The names were then verified and the years were combined  
  + `candy_full_data <- bind_rows(candy_2015_clean, candy_2016_clean, candy_2017_clean)`  

* Missing genders were assigned to "Not gathered"    
  + ...`mutate(gender = coalesce(gender, "Not gathered"))`...  
  
* Where ages were put in the country column (and nothing given in age), the values were moved over to age.  

* Countries were then assigned to "US", "Canada", "UK", and "Other",  
  after converting to lower case and removing punctuation    
  + ...`mutate(country = str_to_lower(country)) %>% `  
    `mutate(country = str_remove_all(country, "[:punct:]"))`...   
  + Those that provided nonsensical country responses have been converted to NA.   
    ...`case_when(`....`str_detect(country, nonsense_pattern) ~ NA,`...  

* Age was extracted, converted to numeric, and then limited to between 4 and 116,  
  all other ages have been converted to NA.  
  + ...`mutate(age = str_extract(age, "[0-9]+"))`...

* Non-candy items have been removed, this includes drugs, bread, comics, wine, and hugs.  
  + ...`case_when(str_detect(candy_type, non_candy_pattern) ~ NA`...  
  
* Some candy items have been recoded based on list of patterns and replacements  
  + eg. anything with "anonymous brown globs...." becomes "mary_janes,  
    "candy_that_is_clearly_just_the_stuff_given_out_for_free_at_restaurants" -> restaurant_candy,  
    etc.


Set up and data format
```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(here)
```

```{r}
candy_data <- read_csv(here("clean_data/candy_data_clean.csv"))
```
[Contents]
# Analysis

## 1. Total number of candy ratings given across the three years. {#Question_1}
Since each row is an individual rating, and all missing ratings have been removed, the number of ratings is equivalent to the number of rows.
```{r question_1}
num_of_ratings <- nrow(candy_data)

cat("The total number of ratings is", format(num_of_ratings, big.mark = ","))
```
[Contents]

## 2. Average age of people who are going out trick or treating. {#Question_2}
This is the average age of individual ids that have put "Yes" for trick_or_treating.
```{r question_2}
average_age <- candy_data %>% 
  group_by(id) %>% 
  filter(trick_or_treating == "Yes") %>% 
  summarise(age = mean(age)) %>% 
  summarise(average_age = mean(age, na.rm = TRUE)) %>% 
  pull()

cat("The average age of people going trick or treating is", round(average_age, 0))
```
[Contents]

## 3. Average age of people who are not going trick or treating? {#Question_3}
This is the average age of individual ids that have put "No" for trick_or_treating.
```{r question_3}
average_age_not_going <- candy_data %>% 
  group_by(id) %>% 
  filter(trick_or_treating == "No") %>% 
  summarise(age = mean(age)) %>% 
  summarise(average_age = mean(age, na.rm = TRUE)) %>% 
  pull()

cat("The average age of people NOT going trick or treating is", round(average_age_not_going, 0))
```
[Contents]

## 4. For each rating which candy bar received the most of these ratings? {#Question_4}
There are three ratings: joy, meh, and despair. 
```{r question_4, warning=FALSE}
candy_data %>%
  group_by(candy_type, rating) %>% 
  summarise(count = n()) %>% # creates a count of the number of each type of ratings each type of candy recieved
  ungroup() %>% # remove candy_type grouping
  group_by(rating) %>% 
  slice_max(count) %>% # max number of ratings for each rating
  select(rating, candy_type, count) # to order the results
```
"Any full sized candy bar" is quite vague.   
This can be removed from the dataset to find a more specific candy with the most joy ratings.  
*(note: it has not been cleaned from the data as this is still an interesting finding that people love candy bars in general)*
```{r question_4_refined, warning=FALSE}
candy_data %>%
  group_by(candy_type, rating) %>% 
  filter(candy_type != "any_full_sized_candy_bar") %>% 
  summarise(count = n()) %>%
  ungroup() %>% 
  group_by(rating) %>% 
  slice_max(count) %>% 
  select(rating, candy_type, count)
```
[Contents]

## 5. How many people rated Starburst as despair? {#Question_5}
```{r question_5}
total_despair <- candy_data %>% 
  filter(str_detect(candy_type, "starburst"), rating == "despair") %>% 
  summarise(n()) %>% 
  pull()

total_ratings <- candy_data %>% 
  filter(str_detect(candy_type, "starburst")) %>% 
  summarise(n()) %>% 
  pull()

percent <- round((total_despair/total_ratings)*100, 0)

cat("The total despair ratings for starbursts were ", format(total_despair, big.mark = ","), ". This was ", percent, "% of the total ratings.", sep = "")
```
[Contents]

### Funtion {#Function}
Numeric Rating System:  
Joy = 1  
Meh = 0  
Despair = -1  

Function to find the most popular candy type within each group (eg. gender, country) excluding "any_full_sized_candy_bar".
```{r most_popular_candy_function}
most_popular_by_category <- function(dataset, category){
  dataset %>% 
        group_by({{category}}, candy_type) %>% 
        filter(candy_type != "any_full_sized_candy_bar") %>% 
        summarise(average_rating = sum(rating_score), number_of_ratings = n()) %>%
        ungroup() %>% 
        group_by({{category}}) %>% 
        slice_max(average_rating)
}
```
[Contents]

## 6. What was the most popular candy bar by this rating system for each gender in the dataset? {#Question_6}
```{r question_6, message=FALSE}
most_popular_by_category(candy_data, gender)
```
[Contents]

## 7. What was the most popular candy bar in each year? {#Question_7}
```{r question_7, message=FALSE}
most_popular_by_category(candy_data, year)
```
[Contents]

## 8. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries? {#Question_8}
```{r question_8, message=FALSE}
most_popular_by_category(candy_data, country)
```
[Contents]

## Extra. Most popular candy by age range {#Extra}
```{r most_popular_by_age, message=FALSE}
candy_data %>% 
  mutate(age_bracket = case_when(
    age <= 10 ~ "10 and under",
    age <= 20 ~ "11-20",
    age <= 30 ~ "21-30",
    age <= 40 ~ "31-40",
    age <= 60 ~ "41-60",
    age > 60 ~ "Over 60"
  )) %>% 
  most_popular_by_category(age_bracket)
```

[Contents]

# Visualisations

## Rating by Country
```{r}
candy_data %>% 
  filter(candy_type %in% c("reese_s_peanut_butter_cups", "skittles", "sourpatch_kids", "licorice", "bonkers", "lollipops"),
         !is.na(country)) %>% 
  group_by(country, candy_type) %>% 
  summarise(rating = mean(rating_score), total_ratings = n()) %>% 
  filter(total_ratings > 10) %>% 
  arrange(desc(rating)) %>% 
  mutate(candy_type = case_when(
    str_detect(candy_type, "reese") ~ "Reese's Peanut Butter Cups",
  .default = str_to_title(str_replace_all(candy_type, "_", " "))),
  country = str_to_upper(country)) %>% 
  ggplot(aes(candy_type, rating, fill = country)) +
  geom_col(position = "dodge") +
  labs(
    title = "Rating of Various Candy Types by Country",
    x = "Candy",
    y = "Average Rating",
    fill = "Country"
  ) +
  scale_y_continuous(limits = c(-1, 1)) +
  coord_flip() +
  theme_classic()

ggsave("candy_ratings_by_country_plot.png")
```
## Survey Responses Received
```{r}
candy_data %>% 
  mutate(year = factor(year, levels = c(2015, 2016, 2017)),
         country = coalesce(country, "Not Given")) %>% 
  group_by(year, country) %>% 
  summarise(survey_results = n_distinct(id)) %>% 
  ggplot(aes(year, survey_results, fill = str_to_upper(country))) +
  geom_col() +
  labs(
    title = "Survey Results Received by Year and Country",
    x = "Year",
    y = "Number of Responses",
    fill = "Country"
  ) +
  theme_classic()

ggsave("survey_results_received.png")
```

