---
title: "Halloween Candy Analysis"
output: html_notebook
---

# Hallowe'en Candy Analysis

## Introduction

## Set up
```{r}
library(tidyverse)
library(here)

candy_data <- read_csv(here("clean_data/candy_data_clean.csv"))
```

## What is the total number of candy ratings given across the three years.
Since each row is an individual rating, and all missing ratings have been removed, the number of ratings is equivalent to the number of rows.
```{r}
num_of_ratings <- nrow(candy_data)

cat("The total number of ratings is", num_of_ratings)
```
## What was the average age of people who are going out trick or treating?
```{r}
average_age <- candy_data %>% 
  group_by(id) %>% 
  filter(trick_or_treating == "Yes") %>% 
  summarise(age = mean(age)) %>% 
  summarise(average_age = mean(age, na.rm = TRUE)) %>% 
  pull()

cat("The average age of people going trick or treating is", round(average_age, 0))
```
## What was the average age of people who are not going trick or treating?
```{r}
average_age_not_going <- candy_data %>% 
  group_by(id) %>% 
  filter(trick_or_treating == "No") %>% 
  summarise(age = mean(age)) %>% 
  summarise(average_age = mean(age, na.rm = TRUE)) %>% 
  pull()

cat("The average age of people not going trick or treating is", round(average_age_not_going, 0))
```
## For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r}
candy_data %>%
  group_by(candy_type, rating) %>% 
  summarise(count = n()) %>% # creates a count of the number of each type of ratings each type of candy recieved
  ungroup() %>% # remove candy_type grouping
  group_by(rating) %>% 
  slice_max(count) %>% 
  select(rating, candy_type, count)
```
"Any full sized candy bar" is quite vague. 
This can be removed from the dataset to find the specific candy with the most joy ratings.
```{r}
candy_data %>%
  group_by(candy_type, rating) %>% 
  filter(candy_type != "any_full_sized_candy_bar") %>% 
  summarise(count = n()) %>%
  ungroup() %>% 
  group_by(rating) %>% 
  slice_max(count) %>% 
  select(rating, candy_type, count)
```

## How many people rated Starburst as despair?
```{r}
candy_data %>% 
  filter(str_detect(candy_type, "starburst"), rating == "despair") %>% 
  summarise(total_despair = n())
```
For the next three questions, count despair as -1, joy as +1, and meh as 0.

```{r}
most_popular_by_category <- function(category){
  #if_else(
    #as.character(category) %in% names(dataframe),
    #(
  
      # dataframe %>% 
      #   group_by(category, candy_type) %>% 
      #   filter(candy_type != "any_full_sized_candy_bar") %>% 
      #   summarise(average_rating = sum(rating_score), number_of_ratings = n()) %>%
      #   ungroup() %>% 
      #   group_by(category) %>% 
      #   slice_max(average_rating)
    
  #),
    #cat("This is not a valid category", names(dataframe)))
  candy_data %>% 
        group_by(category, candy_type) %>% 
        filter(candy_type != "any_full_sized_candy_bar") %>% 
        summarise(average_rating = sum(rating_score), number_of_ratings = n()) %>%
        ungroup() %>% 
        group_by(category) %>% 
        slice_max(average_rating)
}

"gender" %in% names(candy_data)
most_popular_by_category(gender)
```


## What was the most popular candy bar by this rating system for each gender in the dataset?
```{r}
candy_data %>% 
  group_by(gender, candy_type) %>% 
  summarise(average_rating = sum(rating_score)) %>%
  ungroup() %>% 
  group_by(gender) %>% 
  slice_max(average_rating)
```
```{r}
candy_data %>% 
  group_by(gender, candy_type) %>% 
  filter(candy_type != "any_full_sized_candy_bar") %>% 
  summarise(average_rating = sum(rating_score), number_of_ratings = n()) %>%
  ungroup() %>% 
  group_by(gender) %>% 
  slice_max(average_rating)
```
## What was the most popular candy bar in each year?
```{r}
candy_data %>% 
  group_by(year, candy_type) %>% 
  filter(candy_type != "any_full_sized_candy_bar") %>% 
  summarise(average_rating = sum(rating_score), number_of_ratings = n()) %>%
  ungroup() %>% 
  group_by(year) %>% 
  slice_max(average_rating)
```
## What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
```{r}
candy_data %>% 
  group_by(country, candy_type) %>% 
  filter(candy_type != "any_full_sized_candy_bar") %>% 
  summarise(average_rating = sum(rating_score), number_of_ratings = n()) %>%
  ungroup() %>% 
  group_by(country) %>% 
  slice_max(average_rating)
```

